---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: "Riccardo Fusaroli"
date: "3/7/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive patterns of pitch mean and pitch sd in schizophrenia? Report how many papers report quantitative estimates, your method to analyze them, the estimated effect size of the difference (mean effect size and standard error for pitch mean, same for pitch sd) and forest plots representing it. 

2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.

## Tips on the process to follow:

- Download the data on all published articles analyzing pitch in schizophrenia: https://www.dropbox.com/s/pmgw8wzc308so1p/Data.csv?dl=0
- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

- Build a forest plot of the results (forest(model))
 http://rpsychologist.com/d3/cohend/
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

- Now look at the output of rma() and check tau and I2

```{r}
p_load(tidyverse, metafor, lme4, lmerTest)

df = read.csv("Data.csv", sep = ";")

#calculating Cohen's d as well as its variance for pitch mean
df = escalc('SMD', n1i = SampleSizeSchizo, n2i = SampleSizeContros, m1i = PitchMeanSchizo, m2i = PitchMeanControls, sd1i = PitchMeanSchizoSD, sd2i = PitchMeanControlsSD, data = df)

#renaming the yi and vi values (yi = Cohen's D, vi = variance of D)
names(df)[names(df) == 'yi'] = 'yiMean'
names(df)[names(df) == 'vi'] = 'viMean'

#calculating Cohen's d as well as its variance for pitch SD
df = escalc('SMD', n1i = SampleSizeSchizo, n2i = SampleSizeContros, m1i = PitchSDSchizo, m2i = PitchSDControls, sd1i = PitchSDSchizoSD, sd2i = PitchSDControlsSD, data = df )

#renaming the yi and vi values
names(df)[names(df) == 'yi'] = 'yiSD'
names(df)[names(df) == 'vi'] = 'viSD'


#creating models
m1 = lmer(yiMean ~ 1 + (1|Article), weights = 1/viMean, data=df, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m1)

m2 = lmer(yiSD ~ 1 + (1|Article), weights = 1/viSD, data=df, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m2)

#creating metaanalysis optimized models and forest plots
res1 = rma(yiMean, viMean, data = df, slab=Article) 
forest(res1) 

res2 = rma(yiSD, viSD, data = df, slab=Article) 
forest(res2) 

#nothing's significant, everything crosses zero :(
```

Calculating Pitch Mean and SD from assignment 3

```{r}
as3 = read.csv("combinedSchizoData.csv")

TDsub = subset(as3, Diagnosis == "Control")
Schizosub = subset(as3, Diagnosis == "Schizophrenia")

TDmeans = TDsub %>%
  summarise(SampleSizeContros = n(), PitchMeanControls = mean(mean), PitchMeanControlsSD = sd(mean), PitchSDControls = mean(sd), PitchSDControlsSD = sd(sd))


Schizomeans = Schizosub %>%
  summarise(SampleSizeSchizo = n(), PitchMeanSchizo = mean(mean), PitchMeanSchizoSD = sd(mean), PitchSDSchizo = mean(sd), PitchSDSchizoSD = sd(sd))

as3means = cbind(TDmeans, Schizomeans)
as3means$Article = "Bølgen et. al 2017"
as3means$Year = "2017"
```

Adding as3 to the metaanalysis and re-running the analysis.
```{r}
df = read.csv("Data.csv", sep = ";")
df = rbind(df, as3means)

#calculating Cohen's d as well as its variance for pitch mean
df = escalc('SMD', n1i = SampleSizeSchizo, n2i = SampleSizeContros, m1i = PitchMeanSchizo, m2i = PitchMeanControls, sd1i = PitchMeanSchizoSD, sd2i = PitchMeanControlsSD, data = df)

#renaming the yi and vi values (yi = Cohen's D, vi = variance of D)
names(df)[names(df) == 'yi'] = 'yiMean'
names(df)[names(df) == 'vi'] = 'viMean'

#calculating Cohen's d as well as its variance for pitch SD
df = escalc('SMD', n1i = SampleSizeSchizo, n2i = SampleSizeContros, m1i = PitchSDSchizo, m2i = PitchSDControls, sd1i = PitchSDSchizoSD, sd2i = PitchSDControlsSD, data = df )

#renaming the yi and vi values
names(df)[names(df) == 'yi'] = 'yiSD'
names(df)[names(df) == 'vi'] = 'viSD'


#creating models
m1 = lmer(yiMean ~ 1 + (1|Article), weights = 1/viMean, data=df, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m1)

m2 = lmer(yiSD ~ 1 + (1|Article), weights = 1/viSD, data=df, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m2)

#creating metaanalysis optimized models and forest plots
res1 = rma(yiMean, viMean, data = df, slab=Article) 
forest(res1) 
#baaaarely not zero


res2 = rma(yiSD, viSD, data = df, slab=Article) 
forest(res2) 

#tau² and i² 
summary(res1)
summary(res2)


#testing for influential studies 
inf1 = influence(res1) 
print(inf1) 
plot(inf1)
#Our study is influential 

inf2 = influence(res2) 
print(inf2) 
plot(inf2)
#Cohen 2014 is influential


#funnel plooot
plot(res1) #all sorts of plots!!
funnel(res1, main = "Random-Effects Model",xlab = "Standardized Mean Difference") 
#could be worse..


plot(res2) #all sorts of plots!!
funnel(res2, main = "Random-Effects Model",xlab = "Standardized Mean Difference") 
#could be better..


#regtest and ranktest
regtest(res1) 
ranktest(res1) 

regtest(res2) 
ranktest(res2)

```



